---
title: "Curated Cell Data Evaluation"
author: "Kenneth Daily"
date: "`r date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

```{r, include=TRUE, echo=FALSE}
library(tidyverse)
library(knitr)
library(DT)
library(synapser)
library(googlesheets)

url <- 'https://docs.google.com/spreadsheets/d/1DiUuZvZuIFTAEfJlQqK2dcDJFZSsakmFi5Ht5F1TUds/'
curated_cell_table <- 'syn18411380'
syn <- synLogin(silent=TRUE)
```

Get data from the google sheet at `r url`.

```{r}
table_1_gs <- googlesheets::gs_title('NeuroLINCS Table 1')
table_1 <- gs_read(table_1_gs, ws=1)
table_1
```

Get the data from the curated cell table `r curated_cell_table`.

```{r, message=FALSE}
curated_data_obj <- synGet(curated_cell_table)
curatedData <- read.csv(curated_data_obj$path) %>% 
  tibble::as_tibble()
```

Get the list of encoded image masks uploaded to Synapse (syn11688505).

```{r}
imageMasksRes <- synTableQuery("select id,Experiment,Well,PID,parentId from syn11688505")
imageMasks <- imageMasksRes %>% 
  as.data.frame() %>%
  tibble::as.tibble() %>% 
  select(-ROW_ID, -ROW_VERSION, -ROW_ETAG)
```

Get metadata about the tracking datasets (syn11817859).

```{r}
trackingDatasetsRes <- synTableQuery("select * from syn11817859")
trackingDatasets <- trackingDatasetsRes %>% as.data.frame() %>% tibble::as.tibble() %>% 
  select(-ROW_ID, -ROW_VERSION) %>% 
  mutate(CurationFiles=as.logical(CurationFiles),
         CellMasks=as.logical(CellMasks),
         AlignedImages=as.logical(AlignedImages)) %>% 
  filter(!is.na(DatasetName)) %>% 
  arrange(DatasetName)

```

```{r}
trackingDatasets
```

Compare counts with respect to the dataset table for data that has been uploaded or marked as uploaded.

```{r}
trackingDatasets %>% 
  left_join(imageMasks %>% count(Experiment) %>% rename(WellsImageMasks=n)) %>% 
  left_join(curatedData %>% select(Experiment, Well) %>% 
              group_by(Experiment) %>% 
              summarize(WellsInCuratedData=n_distinct(Well))
            ) %>% 
  filter((WellsImageMasks > 0 | CellMasks), (WellsInCuratedData > 0 | CurationFiles)) %>% 
  select(DatasetName, Experiment, Wells=NumberOfWells, WellsImageMasks, CuratedWells, WellsInCuratedData, everything()) %>% 
  arrange(DatasetName)
```


Adjust the timepoints so they all start from 0. This is not necessarily safe if an entire experiment is missing at least one well with the minimum timepoint (e.g., 1 instead of 0).

Filter data that has been manually curated (has a `ObjectTrackID`).

```{r}
minTimePoints <- curatedData %>% 
  group_by(Experiment) %>% 
  summarise(minTimePoint=min(TimePoint))

curatedDataFixed <- curatedData %>% 
  left_join(minTimePoints) %>% 
  mutate(XCoordinate=as.numeric(XCoordinate),
         YCoordinate=as.numeric(YCoordinate),
         TimePointAdjusted=TimePoint-minTimePoint,
         Mistracked=as.logical(Mistracked),
         Live_Cells=as.logical(Live_Cells),
         Out_of_Focus=as.logical(Out_of_Focus),
         Lost_Tracking=as.logical(Lost_Tracking)) %>% 
    filter(!is.na(ObjectTrackID)) #,
    #        !Mistracked, Live_Cells, 
    #        !Out_of_Focus, !Lost_Tracking)
```

I would expect that anything that has been curated for an experiment/well combination (e.g, `ObjectTrackID` is not null) has an image mask file for it. This is the list of missing experiment/well image masks:

```{r}
curatedDataFixed %>% 
  select(Experiment, Well) %>% 
  distinct() %>% 
  anti_join(., imageMasks  %>% select(Experiment, Well) %>% distinct()) %>% 
  group_by(Experiment) %>% 
  summarize(nMissingWells=n_distinct(Well), 
            MissingWells=paste(Well, collapse=","))

```

## Total data points

### Distinct wells and curated objects

Per experiment, how many wells were curated and how many objects total across time points?

```{r, echo=FALSE}
curatedDataFixed %>% 
  left_join(., trackingDatasets, by='Experiment') %>% 
  group_by(DatasetName, Experiment) %>% 
  mutate(`CuratedObjects`=paste0(Well, ObjectTrackID)) %>% 
  summarise_at(vars(Well, TimePoint, CuratedObjects), n_distinct)
```

## Identify duplicates 

These curated objects are in the same well, same timepoint and have the same `ObjectTrackId`.

```{r}
curatedDataFixed %>% 
  count(Experiment, ObjectTrackID, Well, TimePointAdjusted) %>% 
  filter(n > 1)

```

These curated objects are in the same well, same timepoint and have the same `ObjectTrackId` but different coordinates.

```{r message=FALSE, echo=FALSE, include=TRUE, eval=TRUE}
curatedDataFixed %>% 
  group_by(Experiment, ObjectTrackID, Well, TimePointAdjusted) %>% 
  summarize(n=n_distinct(paste(XCoordinate, YCoordinate))) %>% 
  filter(n > 1)
```

These are in the same `Experiment`, `Well`, and `TimePoint` and have the same `ObjectLabelsFound` but multiple `ObjectTrackID`s:

```{r echo=FALSE}
curatedDataFixed %>% 
  group_by(Experiment, Well, TimePoint, ObjectLabelsFound) %>% 
  summarize(n=n_distinct(ObjectTrackID)) %>% 
  filter(n > 1, !is.na(ObjectLabelsFound))
```

```{r}
sessionInfo()
```

