---
title: "Curated Cell Data Evaluation"
author: "Kenneth Daily"
date: "`r date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

```{r, include=TRUE, echo=FALSE}
library(tidyverse)
library(knitr)
library(DT)
library(synapser)

syn <- synLogin(silent=TRUE)

timepointthreshold <- 3
```

Get the data from the curated cell table.

```{r, message=FALSE}
curatedDataRes <- synTableQuery("select * from syn11378063") 
curatedData <- curatedDataRes %>% as.data.frame() %>% tbl_df()
```

Get the list of encoded image masks uploaded to Synapse.

```{r}
imageMasksRes <- synTableQuery("select id,Experiment,Well,PID,parentId from syn11688505")
imageMasks <- imageMasksRes %>% 
  as.data.frame() %>%
  tibble::as.tibble() %>% 
  select(-ROW_ID, -ROW_VERSION, -ROW_ETAG)
```

Get metadata about the tracking datasets.

```{r}
trackingDatasetsRes <- synTableQuery("select * from syn11817859")
trackingDatasets <- trackingDatasetsRes %>% as.data.frame() %>% tibble::as.tibble() %>% 
  select(-ROW_ID, -ROW_VERSION) %>% 
  mutate(CurationFiles=as.logical(CurationFiles),
         CellMasks=as.logical(CellMasks),
         AlignedImages=as.logical(AlignedImages)
         )

```

```{r}
trackingDatasets
```

Compare counts with respect to the dataset table for data that has been uploaded or marked as uploaded.

```{r}
trackingDatasets %>% 
  left_join(imageMasks %>% count(Experiment) %>% rename(WellsImageMasks=n)) %>% 
  left_join(curatedData %>% select(Experiment, Well) %>% 
              group_by(Experiment) %>% 
              summarize(WellsInCuratedData=n_distinct(Well))
            ) %>% 
  filter((WellsImageMasks > 0 | CellMasks), (WellsInCuratedData > 0 | CurationFiles)) %>% 
  select(Experiment, Wells=NumberOfWells, WellsImageMasks, CuratedWells, WellsInCuratedData, everything())
```


Adjust the timepoints so they all start from 0. This is not necessarily safe if an entire experiment is missing at least one well with the minimum timepoint (e.g., 1 instead of 0).

Filter data that has been manually curated (has a `ObjectTrackID`).

```{r}
minTimePoints <- curatedData %>% 
  group_by(Experiment) %>% 
  summarise(minTimePoint=min(TimePoint))

curatedDataFixed <- curatedData %>% 
  select(-ROW_ID, -ROW_VERSION) %>% 
  left_join(minTimePoints) %>% 
  mutate(XCoordinate=as.numeric(XCoordinate),
         YCoordinate=as.numeric(YCoordinate),
         TimePointAdjusted=TimePoint-minTimePoint,
         Mistracked=as.logical(Mistracked),
         Live_Cells=as.logical(Live_Cells),
         Out_of_Focus=as.logical(Out_of_Focus),
         Lost_Tracking=as.logical(Lost_Tracking)) %>% 
    filter(!is.na(ObjectTrackID)) #,
    #        !Mistracked, Live_Cells, 
    #        !Out_of_Focus, !Lost_Tracking)
```

I would expect that anything that has been curated for an experiment/well combination (e.g, `ObjectTrackID` is not null) has an image mask file for it. This is the list of missing experiment/well image masks:

```{r}
curatedDataFixed %>% 
  select(Experiment, Well) %>% 
  distinct() %>% 
  anti_join(., imageMasks  %>% select(Experiment, Well) %>% distinct()) %>% 
  group_by(Experiment) %>% 
  summarize(nMissingWells=n_distinct(Well), 
            MissingWells=paste(Well, collapse=",")) %>% 
  knitr::kable()
```

## Total data points

Per experiment, how many tracked objects and timepoints are there?

```{r}
curatedDataFixed %>% 
  group_by(Experiment) %>% 
  summarize(`Data Points`=n())
```

### Distinct wells and curated objects

Per experiment, how many wells were curated and how many objects total across time points?

```{r, echo=FALSE}
curatedDataFixed %>% group_by(Experiment) %>% 
  mutate(`CuratedObjects`=paste0(Well, ObjectTrackID)) %>% 
  summarise_at(vars(Well, CuratedObjects), n_distinct)
```

## Identify duplicates 

These curated objects are in the same well, same timepoint and have the same `ObjectTrackId` but different coordinates.

```{r message=FALSE, echo=FALSE, include=TRUE, eval=TRUE}
dDuplicate <- curatedDataFixed %>% 
  count(Experiment, ObjectTrackID, Well, TimePointAdjusted) %>% 
  filter(n > 1) %>% 
  arrange(Experiment, Well, ObjectTrackID, TimePointAdjusted)

dDuplicate2 <- dDuplicate %>% 
  left_join(curatedDataFixed) %>% 
  arrange(Experiment, Well, ObjectTrackID, TimePointAdjusted) %>% 
  select(Experiment, Well, ObjectTrackID, TimePointAdjusted, XCoordinate, YCoordinate, Live_Cells, everything())

dDuplicate2
```

These are in the same `Experiment`, `Well`, and `TimePoint` and have the same `ObjectLabelsFound` but multiple `ObjectTrackID`s:

```{r echo=FALSE}
curatedDataFixed %>% 
  group_by(Experiment, Well, TimePoint, ObjectLabelsFound) %>% 
  summarize(n=n_distinct(ObjectTrackID)) %>% 
  filter(n > 1)
```


```{r, message=FALSE, echo=FALSE, include=FALSE, eval=FALSE}
## Identify objects across time
curatedDataFixed %>% 
  count(Experiment, Well, ObjectTrackID, TimePointAdjusted) %>% 
  tidyr::spread(TimePointAdjusted, n) %>% 
  arrange(Experiment, ObjectTrackID, Well) %>% 
  select(Experiment, Well, ObjectTrackID, everything()) %>% 
  DT::datatable()
```

## number of time points distribution

```{r echo=FALSE, message=FALSE}
dtmp2 <- curatedDataFixed %>% 
  group_by(Experiment, ObjectTrackID, Well) %>% 
  summarize(min_time=min(TimePointAdjusted), 
            max_time=max(TimePointAdjusted), 
            ntimepoints=n_distinct(TimePointAdjusted))
```


```{r echo=FALSE, message=FALSE}
p <- ggplot(dtmp2) + geom_histogram(aes(x=ntimepoints), binwidth=1) 
p <- p + facet_wrap(~ Experiment)
p <- p + labs(x="Number of time points", y="Count")
p <- p + theme_bw()
p
```

```{r echo=FALSE, include=FALSE, eval=FALSE}
## Objects with less than `r timepointthreshold` time points:
dtmp2 %>% 
  filter(ntimepoints < timepointthreshold) %>% 
  select(Experiment, ObjectTrackID, Well, ntimepoints)
```

```{r echo=FALSE}
nested_d <- curatedDataFixed %>% 
  group_by(Experiment, ObjectTrackID, Well) %>%
  arrange(TimePointAdjusted) %>% 
  nest()

nested_rle <- nested_d %>% 
  mutate(rlenc=map(data, 
                   function(x) rle(lead(x$TimePointAdjusted) - x$TimePointAdjusted)),
         maxrunlength=map_int(rlenc, function(x) max(x$lengths))) %>% 
  select(Experiment, ObjectTrackID, Well, maxrunlength) %>% 
  arrange(-maxrunlength)
```

## Number of tracked objects with at least `r timepointthreshold` consecutive time points per experiment

```{r message=FALSE}
nested_rle %>% 
  filter(maxrunlength >= timepointthreshold) %>%
  group_by(Experiment) %>% 
  mutate(`CuratedObjects`=paste0(Well, ObjectTrackID)) %>% 
  summarise_at(vars(Well, CuratedObjects), n_distinct)
```

```{r}
sessionInfo()
```

